Do explicitly polarizable force fields give better description of a biomembrane?

### 2023-09-11 manuscriptPOL.tex
revisions of the manuscript text (J.J.M.)

\author{Hanne Antila}
\author{Batuhan Kav}
\affiliation{Forschungszentrum Juelich, Germany}
\email{batuhankav@gmail.com}
\author{Markus S. Miettinen}
\author{O. H. Samuli Ollila}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title[An \textsf{achemso} demo]
%  {A demonstration of the \textsf{achemso} \LaTeX\
%   class\footnote{A footnote for the title}}
\title{NMRLipids VI: Polarizable Force Fields}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\abbreviations{IR,NMR,UV}
%\keywords{American Chemical Society, \LaTeX}

@@

\textcolor{red}{some ideas for the intro/conclusion: 1) amoeba and drude are mostly used for transmembrane protein simulations. https://pubs.acs.org/doi/abs/10.1021/acsnano.1c06443 https://www.sciencedirect.com/science/article/abs/pii/S0959440X2100110X https://pubs.rsc.org/en/content/articlehtml/2021/sc/d1sc01887f etc. Validation/comparison of these simulations with the experiments are done for the proteins, not the lipids. Now since we know that the lipid dynamics and conformations are not very correct with the polarizable force fields, can these affect the protein conformations? is there a way to possibly connect this to the previous nmrlipids jacs paper?\\
}

Conventional classical molecular dynamics (MD) simulations only consider electrostatics as interactions between static point charges, set by molecular dynamics models (force fields) as an input in the beginning of the simulation. The dynamic effects arising from polarizability are ignored or at most enter the simulations in an averaged fashion in the force field parametrization process. However, since decades work~\cite{Thole1981,ando2001stable,Grossfield2003,
lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable } has been underway to make MD simulations explicitly polarizable in the hopes of reaching a more accurate description of the system.

In the case of biomembranes (lipid bilayers), specifically, the molecular (dielectric) environment varies dramatically while crossing the membrane from the water phase across the dipolar/charged headgroup interface to the hydrophobic tail region. Including explicit polarizability in biomembrane modelling has the potential to provide more accurate description of membrane binding, membrane crossing, and behaviour and interactions of molecules residing within membranes, such as membrane protein folding. There certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.

Historically, there has been three main approaches to modelling polarizability: 1) the induced point dipole/multipole model where polarizable point dipoles are placed on chosen sites on a molecule, 2) the classical Drude oscillator model where polarization is modelled by separating a core and a shell charge with a spring which stretches according to the enviroment giving the site a dipole moment, and 3) the electronegativity equilization (fluctuating charge) model where atomic charges are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment. In lipid modelling, the first two have surpassed the third in popularity. All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations.

%To avoid polarizability catastrophe where two closeby polarizable sites feed off from each other and the dipole moments grow infinitely

Considering the growing popularity of both implicitly and explicitly polarizable force fields and the increased computational cost that they require, a systematic benchmark study on polarizable force field quality is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}. Recent publications show that \textit{implicit} inclusion of polarizability improves cation binding simulations when evaluated against NMR data~\cite{Melcr:2018a, melcr2019improved}, which suggest that including polarization explicitly could result in better description of ion binding as well. 
Currently, there are three main polarizable lipid force fields with explicit electronic polarization: CHARMM-Drude~\cite{li2017drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}. 
Here we evaluate the performance of two actively developed polarizable force fields with lipid parameters available, CHARMM-Drude~\cite{li2017drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, against NMR data. We use NMR order parameters to evaluate  structural properties and cation binding, whereas NMR relaxation rates and correlation times are utilized to benchmark the lipid conformational dynamics. Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. Furthermore, we compare the results to the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.

This project has been performed within the NMRLipids collaboration (https://databank.nmrlipids.fi/)~\cite{Databank}. All simulation data, analysis scripts, and results are publicly available in their corresponding links. No plausible nor implausible request is required to download, use, and disseminate the data.

@@

\subsection{Evaluation of lipid bilayer properties simulated with polarizable force fields}

We start by exploring the performance of polarizable force fields using the quality metrics developed in the NMRlipids Databank (https://databank.nmrlipids.fi/)~\cite{Databank}---a quality evaluated databank for lipid simulations. These metrics describe the ability of the models to reproduce the C-H NMR order parameters (quantifying the conformations) and the SAXS data (measuring the electron density variation across the membrane). For the former the molecule is divided into three parts, one containing the headgroup and the glycerol backbone carbons and one for each of the carbonyl tails, and the quality of each is evaluated separately. The quality metrics are summarized in Table~\ref{tab:my_label} and the performance of the force fields is visualized along with the underlying experimental measurables in Fig~\ref{fig:order_parameters}.

One can clearly see that the non-polarizable force fields, CHARMM36 and OPLS3e, consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For older CHARMM-Drude force field, this is likely due to the headgroup and glycerol parametrization having been done targetting absolute average values of the order parameters instead of taking account the potential order parameter forking ( measurably different order parameters for different C--H bods at one carbon) and order parameter sign~\cite{Antila2022}. This was then remedied in the newer version of the model, which also provides better description of the carbon tails. In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.  The AMOEBA force field reproduces the most of the headgroup order parameters well for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in too large forking of oder parameters in this particular carbon. AMOEBA also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}). In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.

The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields for POPC indicate that they are not able to reproduce the SAXS data quite as well as the best non-polarizable force field, OPLS3e. That said, CHARMM-Drude2023, in particular, is clearly better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data. AMOEBA, on the other hand, captures the SAXS profile of POPE very poorly (Table~\ref{tab:my_label} and Fig.~\ref{fig:order_parameters}D top) which is likely connected to the overall description of the lipid tails in the force fields. Note that owing the simulation size dependency of the SAXS data~\cite{Databank} one should focus on the extrema $q_{z}$[\AA$^{-1}$] when comparing the simulated and experimental data. In particular $FFq$ only measures the difference between the first minima of the simulated and experimental curves. Also note, that the method for estimating the charge density of bilayers from the simulations places electrons as point charges at the atom centers in the system and does not consider any redistribution of charge density due to the polarizability. Nevertheless, we expect this results into a good estimate of the charge density for the calculation of the SAXS curve.

\subsection{Evaluation of conformational dynamics of lipids}

We proceed to evaluate (Fig.~\ref{fig:correlation_times}) the structural dynamics of the polarizable models againts NMR $R_{1}$ rates, here measured at a magnetic frequency that captures a window of dynamics around $\sim$0.1-1~ns, and effective correlation times $\tau_{\mathrm{eff}}$ which quantifies a larger range of dynamical processes up to $\sim$1000~ns. We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison. The AMOEBA force field captures the dynamics with respect to both measurables very well,  in par with the best non-polarizable models. In contrast, the dynamics of CHARMM-Drude appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics. The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here. The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow. This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results. The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times. Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding. Note that the experimental data here is from POPC. However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.

Finally, we address the premise that adding polarizability to force fields improves the depiction of ion binding to biomembranes in simulations. For this, we use the concept of "lipid electrometer"~\cite{Catte2016} where the amount of ion binding to the membrane can be quantified by the change in $\alpha$ and $\beta$ order parameters upon increasing salt concentration. These changes in order parameters are depicted in Fig.~\ref{fig:popc_order_parameter_change} whereas Fig.~\ref{fig:ion_density_profiles} gives the density profiles of ions with respect to the bilayer normal. The CHARMM-Drude2017 exhibits strong sodium binding, evidenced both by large change in the order parameters and the ion density profiles. Based on the change in the order parameters, the sodium binding is even stronger than the divalent calcium binding. In strong contrast, experimentally one sees no sodium binding and some binding of calcium ions. Too strong sodium binding in the CHARMM-Drude model has been indicated before for peptides/amino acids~\cite{Ngo2019, Kav2022} and deep-eutectic solvents~\cite{shayestehpour2022ion} along with a too strong calcium interactions~\cite{Tan2022} in line with what we observe here for the lipids.

The faulty sodium ion binding seems to have been somewhat remedied in the CHARMM-Drude2023 force field: the change of the order parameters is diminished to be more in line with experiments and only the some accumulation of sodium ions near the phosphate oxygen is seen in the ion distributions. Based on the change in the order parameters, the calcium binding also appears to be better represented in the updated model. The ion distributions reveal that binding of calcium has moved deeper into the bilayer and there is a stronger formation of ionic double layer by subsequent attraction of chloride into the bilayer. However, this the ion layering is most prominently present in the original CHARMM36 force field, for for sodium and calcium ions. Note, that both for sodium and calcium ion additionin the CHARMM-Drude2023force field reintroduces the forking of order parameters at the $\beta$ carbon in particular. These might also be a sign of poor convergence of the simulations owing to the slow conformational dynamics detected above, which would also not be fully reflected in the calculation of error estimates for the order parameters.

As opposed to the CHARMM36-Drude model, the order parameters of the AMOEBA model are almost completely unaffected by the ion addition, both for NaCl and CaCl$_{2}$. The ion density profiles confirm that very little ions penetrate the membrane region where the lipid phosphates reside. While this is in line with experiments regarding sodium ions, there should be some binding of calcium ions to the membrane. Again, the change in the order parameters indicate the non-polarizable CHARMM36 depicts sodium binding the best out of the models compared here, being slightly more realistic than its most recent polarizable counterpart, CHARMM-Drude2023. Regarding calcium binding, CHARMM36 captures the $\Delta^{\alpha}_{CH}$ about as well as the CHARMM-Drude2023, but the change in the $\beta$ carbon is better represented in the CHARMM-Drude2023 and AMOEBA models. This suggest that some accuracy might have been gained in modelling the divalent ion binding with polarizable models but the advantage is not clear. No force field, polarizable or non-polarizable, is able to model both monovalent and divalent ion binding flawlessly. This is of note, as polarizable force fields are often assumed superior and concequently a popupar choice in modelling, eg, ion channels~\cite{sun2017 jing2019polarizable, klesse2020induced}\todo{I think there is a mistake in the C36 data, firsly, the concentration should probably be 950mM instead of 1000mM and secondly the monovalent anion and cation curves look identical} ~\todo{batuhan: I fxed the anion and cation curves. Yes, C36 data is at 950mM but explicitly adding this as a new color could clutter the figure. Should I just mention this in the caption?}.

@@

\begin{figure}[!hbt]
    \centering
    %\includegraphics{Figures/order_parameters.png}
    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank. The atom labelling is as depicted in Fig.~\ref{fig:molecules}.}
    \label{fig:order_parameters}

@@

\begin{table}[]
    \centering
    \begin{tabular}{c c c c c c c}
        Lipid & Force field & $P^{\mathrm{headgroup}}$ & $P^{sn-1}$ & $P^{sn-2}$ & $FF_{q}$ & APL \\
        \hline
        POPC & OPLS3e       & 0.76 & 0.87 & 0.85 & 0.15 & 66.5\\
        POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
        POPC & CHARMM-Drude & 0.51 & 0.22 & 0.47 & 1.16 & 62.5\\
        POPC & CHARMM-Drude2023 & 0.50 & 0.49 & 0.53 & 0.82 & 64.5\\
        \hline
        POPE & GROMOS-CKP & 0.29 & 0.83 & 0.48 & 0.40 & 59.6\\
        POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
        POPE & CHARMM-Drude2023 & 0.28 & 0.59 & 0.54 & 0.00 & 61.4  \\
        POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
        \hline
        DOPC & AMOEBA & 0.60 & 0.60 & 0.54 & - & 70.2\\
    \end{tabular}
    \caption{NMRlipids databank quality evaluations~\cite{Databank} and area per lipids compared between the best simulations found from the NMRlipids databank, simulations with CHARMM36 force field, and simulations with polarizable force fields. Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}.}
    \label{tab:my_label}
\end{table}

\begin{figure}[!hbt]
	\centering
	\includegraphics{Figures/both_inkscp.pdf}
	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields. Note, top panel axis is logarithmic to visualize the slow Drude dynamics. Experimental values are obtained from Ref.~\citenum{Antila2022rot}. All simulations used here were salt free.}
	\label{fig:correlation_times}

\clearpage

\subsection{Ion binding to membranes in simulations with polarizable force fields}

>>>>

\author{Hanne Antila}
\author{Batuhan Kav}
\affiliation{Forschungszentrum Juelich, Germany}
\email{batuhankav@gmail.com}
\author{Jesper J. Madsen}
\author{Markus S. Miettinen}
\author{O. H. Samuli Ollila}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title[An \textsf{achemso} demo]
%  {A demonstration of the \textsf{achemso} \LaTeX\
%   class\footnote{A footnote for the title}}
%\title{NMRLipids VI: Polarizable Force Fields}
\title{Do explicitly polarizable force fields give better description of a biomembrane?}
%\title{Comparison of polarizable lipid models to high-resolution experimental data reveals room for improvement}
%\title{Benchmarking polarizable lipid force field}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\abbreviations{IR,NMR,UV}
%\keywords{American Chemical Society, \LaTeX}

@@

The dielectric environment in and around a lipid bilayer varies dramatically as you move from the water phase through the interfacial headgroup region to the membrane interior. The water phase highly polar, while the interfacial region contains dipolar headgroups, and the membrane interior is predominantly hydrophobic and nonpolar due to the aliphatic lipid tails.

In conventional molecular dynamics (MD) simulations, electrostatic interactions are described by assigning static point charges to the atoms and molecules at the beginning of the simulation based on the force field definitions. Dynamic effects arising from molecular polarizability are traditionally not explicitly included and are only considered in an averaged fashion in the force field parametrization process when fitting to macroscopic observables. Significant efforts have been dedicated to introduce explicit polarizability into MD simulations, in the hopes of reaching a more accurate representation of the system~\cite{Thole1981,ando2001stable,Grossfield2003,
lamoreux2003,Antila2013 ,Lemkul2016, baker2015polarizable,jing2019polarizable }. For bilayer membrane simulation, including polarizability in the lipids is expected to provide more accurate descriptions, especially in situations pertaining to membrane binding processes, translocation of charged biomolecules across the membrane, and the behavior of molecules residing within membranes, such as membrane proteins~\cite{baker2015polarizable,Lemkul2016, li2017drude,chu2018anionicpolarizable,Lynch21,Chen2021,Melcr:2018a, melcr2019improved,nencini22}. 
%Indeed, even implicit inclusion of polarizability appears to provide a better description of cation binding to membranes~\cite{Melcr:2018a, melcr2019improved}. 
The main lipid force fields with explicit electronic polarization are CHARMM-Drude~\cite{li2017drude, yu2023drude}, AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable}, and CHARMM-Fluctuating Charge (FQ)~\cite{lucas2012charge}. The underlying treatment of polarizability differ among these models in the following way:  1) the classical Drude oscillator (CHARMM-Drude) models polarization by separating a core and a shell charge with a spring which stretches according to the envinronment giving the site a dipole moment~\cite{Lemkul2016},
2) the induced point dipole/multipole (AMOEBA) uses polarizable point dipoles placed on chosen sites of the molecule~\cite{ponder2010current},
and 3) the electronegativity equilization (FQ) employs atomic charges that are no longer constant but can change according to the electronegativities of the molecule atoms and the electric fields from the molecular enviroment~\cite{patel2004charmm}. All of these approaches result in an increasing computational cost, e.g, by introducing new types of interactions, more interactions sites, or by requiring a shorter time step in the simulations. As an alternative approach, electronic continuum correction (ECC) has been proposed as a computational efficient method to implicitly include polarizability by scaling the atom partial charges~\cite{Melcr:2018a,dijon20,Antila2021}.

%Hence, there certainly is room for improvement: efforts from the NMRlipids open science community have shown that the current non-polarizable lipid models cannot correctly reproduce the conformational ensemble~\cite{Botan2015,Antila2019}, conformational dynamics~\cite{Antila2021}, or ion biding~\cite{Catte2016} within MD simulations.

Our previous efforts in benchmarking state-of-the-art non-polarizable force fields have demonstrated that the quality of predictions for important membrane properties greatly vary between different force fields, particularly for lipid headgroup conformational ensembles and ion binding affinities~\cite{Botan2015,Catte2016,Antila2019,bacle21,Antila2021,Antila2022,Databank}. This emphasizes the critical importance of rigorous comparative quality evaluations of simulations performed using different force fields, which helps to estimate the reliability of simulation results and select the best models for specific applications. This is particularly important when employing polarizable force fields, as they offer the potential to capture a broader range of physical phenomena pertinent to enhancing our understanding of in particular polar membrane regions, at an increased computational cost. Unfortunately, systematic benchmark studies assessing the quality of polarizable force fields are,  
%is in great need. The few published studies reviewing the state of polarizable force fields for biomolecule simulations did not go beyond citing results from the existing studies
to our knowledge, entirely missing from the field~\cite{inakollu2020polarisable,jing2019polarizable,baker2015polarizable}. Previous simulations using polarizable lipid force fields center around protein systems, with experimental validation (if performed at all) fovcusing on the properties of the proteins~\cite{jing2019polarizable, Lynch21}. Assessing the realism of the lipid models against experimental evidence from biophysical measurements is crucial since these membrane properties are foundational to more complicated applications, such as studies of transmembrane proteins. For example, it appears to be important that a wide variety of conformations is available for lipid headgroups in the model for realistic protein-lipid binding~\cite{bacle21}.

Here we set out to assess the quality of available polarizable force fields, on par with previous evaluations of non-polarizable force fields, using the resources given by the NMRLipids open collaboration framework (https://databank.nmrlipids.fi/). Specifically, we evaluated the quality of POPC, DOPC and POPE lipid bilayer simulations with CHARMM-Drude~\cite{li2017drude,yu2023drude} and AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters against experimental NMR and small angle X-ray scattering (SAXS) data. These two force fields are increasingly utilized in biomolecular simulations and have lipid parameters available for the selected systems with corresponding experimental data available from the NMRlipids databank~\cite{Databank}. Following the quality metrics defined in the NMRlipids project, structural properties and cation binding were evaluated against NMR order parameters and SAXS form factors~\cite{Botan2015,ollila16,Catte2016,Databank}, while lipid conformational dynamics was evaluated in comparison to data from NMR spin relaxation rate experiments~\cite{ferreira15,Antila2021}. 
%Small angle scattering (SAXS) data is used to estimate the description of electron density profile of the membrane. 
Furthermore, we compare the results to those obtained from the best performing non-polarizable simulations from the NMRlipids Databank\cite{Databank}.

 %n lipid modelling, the first two have surpassed the third in popularity.

%All simulation data, analysis scripts, and results are publicly available in their corresponding links. No plausible nor implausible request is required to download, use, and disseminate the data.

@@

\subsection{Evaluation of lipid bilayer structural properties}

To evaluate the structural properties of lipid bilayers in simulations with polarizable force fields, we simulated POPC and POPE lipid bilayers with CHARMM-Drude2017~\cite{li2017drude} and CHARMM-Drude2023~\cite{yu2023drude} parameters, and POPE and DOPC bilayers with AMOEBA~\cite{chu2018anionicpolarizable,chu2018polarizable} parameters. These systems were selected based on simultaneous availability of force field parameters, and experimental data~\cite{Databank}. After the data production, simulation trajectories were added into the NMRlipids Databank (https://github.com/NMRLipids/Databank/) and the NMRlipids quality metrics was used to evaluate each simulation against experiments~\cite{Databank}. This metric measures the quality of conformational ensemble of individual lipids in two aspects: first, it evaluates agreement with C-H bond order parameters (from NMR), and second, it examines consistency of membrane dimensions compared to small angle X-ray scattering (SAXS) data~\cite{Databank,ollila16}. 
%(measuring the electron density variation across the membrane). 
The former is divided into three parts that separately describe the average quality of headgroup, glycerol backbone and acyl chain qualities as described in the NMRlipids databank~\cite{Databank}. While order parameters primarily reflect lipid conformations, acyl chain order parameters are also a good proxy for membrane packing since a smaller area per lipid tends to increase the magnitude of the order parameters~\cite{Databank}. 
%one for each of the carbonyl tails, and the quality of each is evaluated separately.

@@

\begin{table}[]
    \centering
    \begin{tabular}{c c c c c c c}
        Lipid & Force field & $P^{\mathrm{headgroup}}$ & $P^{sn-1}$ & $P^{sn-2}$ & $FF_{q}$ & APL \\
        \hline
        POPC & OPLS3e       & 0.76 & 0.87 & 0.85 & 0.15 & 66.5\\
        POPC & CHARMM36     & 0.70 & 0.54 & 0.69 & 1.16 & 65.0\\
        POPC & CHARMM-Drude & 0.52 & 0.29 & 0.53 & 1.06 & 62.5\\
        POPC & CHARMM-Drude2023 & 0.63 & 0.60 & 0.57 & 0.96 & 64.5\\
        \hline
        POPE & GROMOS-CKP & 0.29 & 0.83 & 0.48 & 0.40 & 59.6\\
        POPE & CHARMM36   & 0.54 & 0.52 & 0.27 & 1.30 & 57.2 \\
        POPE & CHARMM-Drude & 0.06 & 0.53 & 0.27 & 0.80 & 56.6  \\
        POPE & CHARMM-Drude2023 & 0.28 & 0.59 & 0.54 & 0.00 & 61.4  \\
        POPE & AMOEBA & 0.21 & 0.10 & 0.23 & 3.80 & 66.9\\
        \hline
        DOPC & AMOEBA & 0.60 & 0.60 & 0.54 & - & 70.2\\
    \end{tabular}
    \caption{NMRlipids databank quality metrics~\cite{Databank} and area per lipids (APL) compared between the best simulations found from the NMRlipids databank (OPLS3e for POPC and GROMOS-CKP for POPE), simulations with CHARMM36 force field, and simulations with polarizable force fields. Order  parameter quality measures, $P^{\mathrm{headgroup}}$, $P^{sn-1}$, and $P^{sn-2}$, present the average probability of order parameters within a given segment to agree with experiments (larger probabilities mean higher quality). Form factor quality measure, $FF_{q}$, presents the difference in essential features between simulated and experimental form factors (smaller values indicate higher quality). Experimental estimations for areas per lipid are POPC: 64.3$\pm1$\AA$^{2}$\cite{kucerka2011}, DOPC: 67.5$\pm1$\AA$^{2}$\cite{kucerka2008}, and POPE: 56.7$\pm$3\AA$^{2}$\cite{Rickeard2020}.}
    \label{tab:my_label}
\end{table}

\begin{figure}[!p]
    \centering
    %\includegraphics{Figures/order_parameters.png}
    \includegraphics[width=0.8\textwidth]{Figures/quality.pdf}
    \caption{X-ray scattering form factors, and C-H bond order parameters for headgroup and glycerol backbone, and acyl chains (from top to bottom) compared between simulations and experiments using the NMRlipids databank. The atom labelling is as depicted in Fig.~\ref{fig:molecules}. Experimental data in this figure are originally reported in Refs.~\citenum{ferreira13,bacle21,Databank,NMRlipidsIII,kucerka2015}. For CHARMM-Drude2023 simulations we selected representative replicas among three available simulations, for example for differences between replicas for POPC see Fig.~\ref{fig:order_parameters_replicas}.}
    \label{fig:order_parameters}

@@

Direct comparisons between simulated and experimental data are visualized in Fig~\ref{fig:order_parameters} and the resulting quality metrics are shown in Table~\ref{tab:my_label}. Similarly to the non-polarizable CHARMM36 simulations with Gromacs, CHARMM-Drude simulation predicts slightly too packed membrane with excessively negative order parameters with respect to experiments and simulations with the highest quality in the NMRlipids databank (OPLS3e)~\cite{Databank}. However,
%consistently outperform the polarizable force fields with respect to the NMR measurables for all parts of the molecule (higher $P$-values). For older CHARMM-Drude force field, 
the quality of headgroup conformations in the CHARMM-Drude simulations is worse than in its non-polarizable counterpart. This is likely because headgroup and glycerol backbone parameters were optimized to reproduce absolute average values of the experimentally determined order parameters, instead of taking into account the order parameter forking (meaning measurably different order parameters for different C--H bonds at a single carbon atom) and order parameter sign~\cite{Antila2022}. A better description of the headgroup and glycerol backbone is provided by CHARMM36-Drude2023, yet its quality is still below that offered by the non-polarizable CHARMM36 parameters (see also discussion about conformational dynamics below). Also the quality of membrane packing and acyl chain order are improved in CHARMM36-Drude2023 compared to the earlier version but again, it did not perform at a quality comparable to that of the best available simulations.
%also provides better description of the carbon tails. 
%In fact, for POPE sn-2 tail the CHARMM-Drude2023 gives the best description of the models investigated here.

 AMOEBA simulations capture the headgroup and glycerol backbone order parameters reasonably well with the exception of g$_1$,
%ll for both DOPC and POPE but vastly overestimates one of the g$_3$ values, resulting in 
where forking of order parameters is unacceptably large. However, experimentally observed low order parameters at double bond region in both DOPC acyl chains and {\it sn}-2 chain of POPE are not captured. These low order parameters signal an important mechanism through which acyl chain double bonds affect membrane properties~\cite{ollila07} and are well reproduced in all state-of-the art MD simulation parameters~\cite{ollila16}. Furthermore, AMOEBA parameters substantially overestimate the area per lipid in POPE simulations, leading to too disordered acyl chains. Similar issues are evident also in membrane data presented in the recent publication for the AMOEBA cholesterol model~\cite{Li23chol}.
%also clearly struggles to depict the tail conformations, in particular in the double bond region of DOPC (Fig.~\ref{fig:order_parameters}). 
%parameters do not capture the  
%In the tail regions, the CHARMM-Drude2023 appears to be the best out of the polarizable models.

The unsatisfactory description of the lipid tail region and the area per lipid is further reflected in the inability of AMOEBA to capture the SAXS curves. Thus, we can conclude that our simulations with the AMOEBA parameters did not reproduce essential
membrane properties. The higher $FFq$ values (Table~\ref{tab:my_label}) of the polarizable force fields in general indicate that they are not able confer consistency with the SAXS data as well as the best non-polarizable force fields (OPLS3e). That said, CHARMM-Drude2023 is on par with or better than its non-polarizable counterpart CHARMM36 and the best out of all the models here in capturing the POPE SAXS data, benefiting from the improvements made in the description of acyl chains.

\subsection{Evaluation of conformational dynamics of lipids}
C-H bond order parameters are sensitive to the conformational ensemble but the correspondence is not unique. Instead, they essentially describe the average of the distributions and do not contain information of the dynamics of the confomational sampling. Therefore, a simulation that reproduces the order parameters has an ensemble that is only potentially correct, and even the correct ensemble may not have the experimentally observed dynamics. To further elucidate this aspect in the case of polarizable force fields, we compared NMR $^{13}$C spin relaxation $R_{1}$ rates and C-H bond effective correlation times $\tau_{\mathrm{eff}}$ with experiments~\cite{ferreira15} and best non-polarizable simulations from our previous study~\cite{Antila2021} for simulations with PC headgroups and the glycerol backbone in Fig.~\ref{fig:correlation_times}. This focus was chosen based on the availability of both experimental data and polarizable simulations. $R_{1}$ rates measured with typical magnetic field strengths are sensitive to rotational dynamics of C-H bonds with the timescales around $\sim$0.1-1~ns, while effective correlation times respond to a larger range of dynamical processes up to $\sim$1000~ns~\cite{ferreira15}. The effective correlation time can be interpreted as an average measure for how fast molecular conformations sample the phase space that leads to the average order parameters.

The effective correlation times in CHARMM-Drude2017 and CHARMM-Drude2023 simulations are approximately two  and one orders of magnitude slower, respectively, than the values suggested by experiments and best available simulations (Fig.~\ref{fig:correlation_times}). This indicates that not only are the simulations with this polarizable force field computationally costlier for equivalent lengths of trajectory, but one would also have to simulate longer to obtain converged results. The non-polarizable counterpart of the Drude models, CHARMM36, exhibits much more realistic, faster effective correlation times. Also the dynamics in the fast range ($R_{1}$ rates), are on average slightly more realistic in the non-polarizable CHARMM36 compared to both of the polarizable counterparts. Interestingly, CHARMM-Drude models have been reported to have slower water hydrogen bonding dynamics around amino acids compared to their non-polarizable counterpart,~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding.

In contrast to the Drude-based models investigated here, the $R_1$ spin relaxation times and effective correlation times from DOPC simulations with AMOEBA force field are close to experimental data from POPC and on par with the best non-polarizable models (Slipids and CHARMM36). The small difference in acyl chain composition (DOPC vs. POPC) is not expected to strongly affect headgroup dynamics due to effective decoupling between hydrophilic and hydrophobic membrane regions~\cite{Antila2022rot, Klauda08}.
%appears to be faulty in the fast range ($R_{1}$ rates) and, furthermore, $\tau_{\mathrm{eff}}$ values reveal vastly too slow overall dynamics. The updated CHARMM-Drude2023 does not significantly improve the dynamics at the fast range investigated here. The effective correlation times of the newer Drude model are improved but still an order of magnitude too slow. This indicates that not only are the simulations with this polarizable force field computationally costly but one would have to simulate longer to obtain converged results. The non-polarizable counterpart of the Drude model, CHARMM36, exhibits much more realistic, faster effective correlation times. 
%Interestingly, it has also been observed that CHARMM-Drude models have slower dynamic of water hydrogen bonding around amino acids compared to their non-polarizable counterpart~\cite{Ngo2019} which might align with an overall slower dynamics of the model in addition to enhanced water binding. 
%Note that the experimental data here is from POPC. However, chemically identical the headgroup region the dynamics is assumed to be practically identical in DOPC, due to the decoupling of the dynamics between the headgroup and the glycerol regions~\cite{Antila2022rot, Klauda08}.

%We include the same measurables calculated from non-polarizable CHARMM36 and SLipids force fields---the best MD force fields regarding the dynamics~\cite{Antila2021}---as a point of comparison.

\begin{figure}[!hbt]
	\centering
	\includegraphics{Figures/both_inkscp.pdf}
	\caption{Effective correlation times (top) and R$_1$ rates (bottom) for the polarizable models and best performing non-polarizable force fields. Note, top panel axis is logarithmic to visualize the slow Drude dynamics. Experimental values are obtained from Ref.~\citenum{Antila2022rot}. All simulations used here were salt free.}
	\label{fig:correlation_times}
\end{figure}
